---
title: "SIG03 T7oBC 6RT Mouse 10x Seurat Processing"
author: "Eric Y. Wang"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_float: true
  github_document:
    toc: true
    html_preview: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(Seurat)
library(hdf5r)
library(ggplot2)
library(clustree)
library(cowplot)
library(scDblFinder)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("functions/plotting_fxns.R")
source("functions/scRNA_seq_analysis_functions.R")
theme_set(theme_Publication())
```

## [**Import data**]{.underline}

```{r}
dataRaw <- Read10X_h5("C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/cellranger_outs/SIG03_GEX_HTO_T7oBC_6RT_mouse/filtered_feature_bc_matrix.h5")
names(dataRaw)
```

```{r}
# identify joint cells between BC and GEX dataset
jointCells <- intersect(colnames(dataRaw$`Gene Expression`),
                        colnames(dataRaw$`Antibody Capture`))
jointCells <- intersect(jointCells, colnames(dataRaw$`Custom`))

# subset each dataset to include joint cells
dataGEX <- dataRaw$`Gene Expression`[,jointCells]
dataHTO <- dataRaw$`Antibody Capture`[,jointCells]
dataBC <- dataRaw$`Custom`[,jointCells]
```

Remove genes that are expressed by less than 1% of cells (assuming 30% dropout
rate). Set this level based on the smallest size cluster that you want to be
detectable in the final dataset.

```{r}
keepGenes <- Matrix::rowSums(dataGEX > 0) >= ncol(dataGEX)*0.01*0.3
dataGEX <- dataGEX[keepGenes,]
```

```{r}
# load GEX counts and add celltag counts into seurat object
data <- CreateSeuratObject(counts = dataGEX, project = "SIG03_mouse")
data[["HTO"]] <- CreateAssayObject(counts = dataHTO)
data[["BC"]] <- CreateAssayObject(counts = dataBC)

rm(dataGEX)
rm(dataHTO)
rm(dataBC)
```

## [**Doublet Processing**]{.underline}

### HTO processing

```{r}
data <- NormalizeData(data, assay = "HTO", normalization.method = "CLR")
data <- HTODemux(data, assay = "HTO", seed = 314)
```

```{r, fig.width=6, fig.height=5}
# Calculate a UMAP embedding of the HTO data
DefaultAssay(data) <- "HTO"
data <- ScaleData(data, features = rownames(data),
    verbose = FALSE)
data <- RunPCA(data, features = rownames(data), approx = FALSE, verbose = F)
data <- RunTSNE(data, dims = 1:nrow(data), verbose = F)
```

```{r, fig.width=12, fig.height=8}
# create heatmap of HTO expression
p1 <- HTOHeatmap(data, assay = "HTO")
Idents(data) <- "HTO_classification.global"
# compare UMI number for singlets, doublets, negatives
p2 <- VlnPlot(data, features = "nCount_RNA", pt.size = 0.1, log = TRUE, alpha = 0.5)
# plot HTO data on tSNE
p3 <- DimPlot(data)

(p3+p2)/p1
```

```{r, fig.width=12, fig.height=16}
# vsiualize ridge plots for HTO
Idents(data) <- "HTO_maxID"
RidgePlot(data, assay = "HTO", features = rownames(data[["HTO"]]), ncol = 3)
```

### scDblFinder

```{r}
DefaultAssay(data) <- "RNA"
# create logical vector of doublet or not
doubletHTO <- data$HTO_classification.global == "Doublet"

# run scDblFinder with known doublets from HTO as input
# use GEM-X doublet rate of 0.4% per 1000 cells
sceData <- scDblFinder(GetAssayData(data, layer="RNA", slot="counts"),knownDoublets = doubletHTO, knownUse = "discard", dbr = 0.004)
# port the resulting scores back to the Seurat object:
data$scDblFinder.score <- sceData$scDblFinder.score
data$scDblFinder.class <- sceData$scDblFinder.class
```

```{r}
table(HTO=data$HTO_classification.global, call=sceData$scDblFinder.class)
```

### Subset data to singlets based on HTO

```{r}
data@meta.data %>%
  group_by(HTO_classification.global) %>%
  summarize(num_cells = n())

Idents(data) <- "HTO_classification.global"
dataSub <- subset(data, idents = "Singlet")
```

```{r}
dataSub@meta.data %>%
  group_by(hash.ID) %>%
  summarize(num_cells = n()) %>%
  mutate(percent = num_cells/sum(num_cells)*100)
```

## [**Cell QC Analysis**]{.underline}

### Calculate QC metrics

```{r}
DefaultAssay(dataSub) <- "RNA"
# calculate percent mitochondrial reads and ratio 
dataSub[["percent.mt"]] <- PercentageFeatureSet(dataSub, pattern = "^mt-")
dataSub[["percent.ribo"]] <- PercentageFeatureSet(dataSub, pattern = "^Rpl|^Rps")
dataSub[["mitoRatio"]] <- dataSub@meta.data$percent.mt / 100
dataSub[["log10GenesPerUMI"]] <- log10(dataSub$nFeature_RNA)/log10(dataSub$nCount_RNA)
dataSub[["log10nFeature_RNA"]] <- log10(dataSub$nFeature_RNA)
dataSub[["log10nCount_RNA"]] <- log10(dataSub$nCount_RNA)
metadata <- dataSub@meta.data
```

### Visualize QC distributions

```{r, fig.width=15, fig.height=7}
# enter QC values
qcValues <- c(nFeature_RNA = 400, nCount_RNA = 500, percent.mt = 5)

# quantify how many cells there are
paste0("There are ", ncol(dataSub), " cells")

# Visualize the number UMIs/transcripts per cell
p2 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	ylab("Cell density") +
  	geom_vline(xintercept = qcValues[2]) +
    NoLegend()

# Visualize the distribution of genes detected per cell via histogram
p3 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=nFeature_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	geom_vline(xintercept = qcValues[1]) +
    NoLegend()

# Visualize the correlation between genes detected and number of UMIs and determine whether strong presence of cells with low numbers of genes/UMIs
p4 <- metadata %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
  	scale_colour_gradient(low = "gray90", high = "black") +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	geom_vline(xintercept = qcValues[2]) +
  	geom_hline(yintercept = qcValues[1])


# Visualize the distribution of mitochondrial gene expression detected per cell
p5 <- metadata %>% 
  	ggplot(aes(color=orig.ident, x=percent.mt, fill=orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	geom_vline(xintercept = qcValues[3]) +
    xlim(0,10) +
    NoLegend()

# Visualize the overall complexity of the gene expression by visualizing the genes detected per UMI
p6 <- metadata %>%
  	ggplot(aes(x=log10GenesPerUMI, color = orig.ident, fill=orig.ident)) +
  	geom_density(alpha = 0.2) +
  	geom_vline(xintercept = 0.8) +
    NoLegend()

plot_grid(p2,p3,p4,p5,p6, ncol = 3)
```

```{r, fig.height=8, fig.width=16}
p1 <- metadata %>% 
  ggplot(aes(color=hash.ID, x=nFeature_RNA, fill=hash.ID)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	geom_vline(xintercept = qcValues[1])
    
p2 <- metadata %>% 
  ggplot(aes(color=hash.ID, x=nCount_RNA, fill=hash.ID)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	geom_vline(xintercept = qcValues[2])

p3 <- metadata %>% 
  ggplot(aes(color=hash.ID, x=percent.mt, fill=hash.ID)) + 
  	geom_density(alpha = 0.2) + 
  	geom_vline(xintercept = qcValues[3]) +
    xlim(0,10)

p4 <- metadata %>%
  ggplot(aes(x=log10GenesPerUMI, color = hash.ID, fill=hash.ID)) +
  	geom_density(alpha = 0.2)

plot_grid(p1,p2,p3,p4, ncol = 2)
```

### Subset data based on QC metrics

```{r}
dataSubQC <- subset(dataSub, subset = (nFeature_RNA >= qcValues[1]) & (nCount_RNA >=qcValues[2]) & (percent.mt < qcValues[3]))

# quantify how many cells there are
paste0("There are ", ncol(dataSubQC), " cells after QC")
paste0("There were ", ncol(dataSub) - ncol(dataSubQC), " cells removed in QC")
```

## [**Score for sources of variation**]{.underline}

### Cell cycle variation

```{r, echo=FALSE}
DefaultAssay(dataSubQC) <- "RNA"
dataSubQC <- NormalizeData(dataSubQC)
# Identify the most variable genes
dataSubQC <- FindVariableFeatures(dataSubQC, 
                     selection.method = "vst",
                     nfeatures = 2000, 
                     verbose = FALSE)
# Scale the counts
dataSubQC <- ScaleData(dataSubQC)

cellCycleGenes <- read_csv("C:/Users/Eric/My Drive/Lab/datasets/gene_sets/cell_cycle_genes_seurat_mouse.csv")
# cellCycleGenes <- read_csv("/Users/wange7/Library/CloudStorage/GoogleDrive-ericwang314@gmail.com/My Drive/Lab/datasets/gene_sets/cell_cycle_genes_seurat_mouse.csv")

dataSubQC <- CellCycleScoring(dataSubQC, s.features = cellCycleGenes$s_genes,
                            g2m.features = cellCycleGenes$g2m_genes)
```

```{r, fig.width=8, fig.height=4}
# Plot the PCA colored by cell cycle phase
dataSubQC <- RunPCA(dataSubQC, verbose = FALSE)
DimPlot(dataSubQC,
        reduction = "pca",
        group.by= "Phase",
        split.by = "Phase")
```

### Mitochondrial variation

```{r, fig.width = 12, fig.height=4}
# Check quartile values
mitoQuart <- summary(dataSubQC@meta.data$mitoRatio)

# Turn mitoRatio into categorical factor vector based on quartile values
dataSubQC@meta.data$mitoFr <- cut(dataSubQC@meta.data$mitoRatio, 
                   breaks=c(-Inf, mitoQuart[2], mitoQuart[3], mitoQuart[5], Inf), 
                   labels=c("Low","Medium","Medium high", "High"))

DimPlot(dataSubQC,
        reduction = "pca",
        group.by= "mitoFr",
        split.by = "mitoFr")
```

## [**Data normalization/scaling**]{.underline}

### SCTransform

```{r}
options(future.globals.maxSize= 2000*1024^2)
# perform SCT normalization and find variable features
dataSubQC <- SCTransform(dataSubQC, vst.flavor = "v2",
                       vars.to.regress = c("Phase","mitoRatio"),
                       variable.features.n = 3100)

# remove TCR genes from variable features
featureVar <- VariableFeatures(dataSubQC, assay = "SCT")
TCRgenes <- grep("^Tra[vj]|^Trb[vdj]",featureVar)
BCRgenes <- grep("^Igk[vc]|^Igl[vc]|^Ighv",featureVar)
featureVar[TCRgenes]
featureVar[BCRgenes]

VariableFeatures(dataSubQC, assay = "SCT") <- featureVar[-c(TCRgenes,BCRgenes)]

# scale SCT variable features in RNA assay
DefaultAssay(dataSubQC) <- "RNA"
dataSubQC <- ScaleData(dataSubQC, features = featureVar[-c(TCRgenes,BCRgenes)])
DefaultAssay(dataSubQC) <- "SCT"
```

### BC normalization

```{r}
dataSubQC <- NormalizeData(dataSubQC, assay = "BC", normalization.method = "CLR")
```

## [Sample Subsetting]{.underline}

```{r}
dataSubQC$sample_group <- gsub("^(\\w+)-.*$", "\\1", dataSubQC$hash.ID)

dataCD4 <- subset(dataSubQC, subset = sample_group != "splenocytes")
dataSplenocytes <- subset(dataSubQC, subset = sample_group == "splenocytes")
```

## [Dim Reduction CD4]{.underline}

```{r}
# Find variable features and exclude TCR genes
dataCD4 <- FindVariableFeatures(dataCD4,
                                selection.method = "vst",
                                nfeatures = 3100)
featureVar <- VariableFeatures(dataCD4)
VariableFeatures(dataCD4, assay = "SCT") <- featureVar[-c(TCRgenes)]


# run PCA using TCR/BCR excluded Variable Features
dataCD4 <- RunPCA(dataCD4, npcs = 60, verbose = FALSE)
ElbowPlot(dataCD4, ndims = 60) +
  ggtitle("dataCD4 Elbow") +
  scale_x_continuous(breaks = seq(0,60,3))
```

```{r}
dimValues <- 50
# run umap based on dimensionality from elbow plot
dataCD4 <- RunUMAP(dataCD4, dims = 1:dimValues, verbose = F) %>%
  FindNeighbors(dims = 1:dimValues, verbose = FALSE) %>%
  FindClusters(resolution = seq(0,1.2,0.1),
               algorithm = 1, verbose = F)
```

```{r, echo=FALSE, fig.height=10, fig.width=6}
clustree(dataCD4, prefix = "SCT_snn_res.") +
  ggtitle("dataCD4 Clustree")
```

```{r, fig.width=12, fig.height=5}
dataCD4$seurat_clusters <- dataCD4$SCT_snn_res.0.7
Idents(dataCD4) <- dataCD4$seurat_clusters
p1 <- DimPlot(dataCD4) + theme(aspect.ratio = 1)
p2 <- DimPlot(dataCD4, group.by = "hash.ID") + theme(aspect.ratio = 1)
p1+p2
```

```{r, fig.height=25, fig.width=10}
# perform DEG analysis on SCT data
dataCD4 <- PrepSCTFindMarkers(dataCD4)
clusMarkers <- FindAllMarkers(dataCD4, only.pos = T, assay = "SCT")
clusMarkers <- as_tibble(clusMarkers)

sigMarkers <- clusMarkers %>%
  dplyr::filter(avg_log2FC > 1 & p_val_adj < 0.1) %>%
  arrange(p_val_adj)

top10 <- sigMarkers %>%
    group_by(cluster) %>%
    slice_head(n = 10) %>%
    ungroup()

top10
```

Cluster 5, 14, 15 seem to be contaminating B cells. Cluster 12,13 seems to be
proliferating cells? I will remove them in the final dataset. Cluster 10 is
weird (associated with ribosomal RNA contamination?

```{r}
dataCD4Sub <- subset(dataCD4, invert = TRUE, subset = seurat_clusters %in% c("5","14","15","13","12"))

DimPlot(dataCD4Sub) + theme(aspect.ratio = 1)
```

## [Dim Reduction Splenocytes]{.underline}

```{r}
# Find variable features and exclude TCR genes
dataSplenocytes <- FindVariableFeatures(dataSplenocytes,
                                selection.method = "vst",
                                nfeatures = 3100)
featureVar <- VariableFeatures(dataSplenocytes)
VariableFeatures(dataSplenocytes, assay = "SCT") <- featureVar[-c(TCRgenes,BCRgenes)]

dataSplenocytes <- RunPCA(dataSplenocytes, npcs = 60, verbose = FALSE)
ElbowPlot(dataSplenocytes, ndims = 60) +
  ggtitle("dataSplenocytes Elbow") +
  scale_x_continuous(breaks = seq(0,60,3))
```

```{r}
dimValues <- 50
# run umap based on dimensionality from elbow plot
dataSplenocytes <- RunUMAP(dataSplenocytes, dims = 1:dimValues, verbose = F) %>%
  FindNeighbors(dims = 1:dimValues, verbose = FALSE) %>%
  FindClusters(resolution = seq(0,2,0.1),
               algorithm = 1, verbose = F)
```

```{r, echo=FALSE, fig.height=10, fig.width=6}
clustree(dataSplenocytes, prefix = "SCT_snn_res.") +
  ggtitle("dataSplenocytes Clustree")
```

```{r}
dataSplenocytes$seurat_clusters <- dataSplenocytes$SCT_snn_res.1.2
Idents(dataSplenocytes) <- dataSplenocytes$seurat_clusters
DimPlot(dataSplenocytes) + theme(aspect.ratio = 1)
```

```{r}
# perform DEG analysis on SCT data
dataSplenocytes <- PrepSCTFindMarkers(dataSplenocytes)
clusMarkers <- FindAllMarkers(dataSplenocytes, only.pos = T, assay = "SCT")
clusMarkers <- as_tibble(clusMarkers)

sigMarkers <- clusMarkers %>%
  dplyr::filter(avg_log2FC > 1 & p_val_adj < 0.1) %>%
  arrange(p_val_adj)

top10 <- sigMarkers %>%
    group_by(cluster) %>%
    slice_head(n = 10) %>%
    ungroup()

top10
```

```{r}
metadata <- dataSplenocytes@meta.data %>%
  as_tibble() %>%
  mutate(clusters_anno = case_when(seurat_clusters %in% c("0","1","2","3","7") ~ "Bcells",
                                   seurat_clusters %in% c("4","8") ~ "Tcell",
                                   seurat_clusters %in% c("5") ~ "5",
                                   seurat_clusters == "6" ~ "NKcell",
                                   seurat_clusters == '10' ~ "10",
                                   seurat_clusters == "9" ~ "Macrophage",
                                   seurat_clusters == "11" ~ "Neutrophil"))

dataSplenocytes$clusters_anno = metadata$clusters_anno
```

```{r}
Idents(dataSplenocytes) <- "clusters_anno"
DimPlot(dataSplenocytes) + theme(aspect.ratio = 1)
```

## [**Save RDS**]{.underline}

```{r}
saveRDS(dataCD4Sub, "C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/seurat_outs/SIG03_mouse_T7oBC6RT_CD4.rds")
saveRDS(dataSplenocytes, "C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/seurat_outs/SIG03_mouse_T7oBC6RT_splenocytes.rds")
```
