---
title: "SIG03 oBC Direct Amp Barcode Analysis"
author: "Eric Y. Wang"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    html_preview: false
  html_notebook:
    toc: true
    toc_float: true
---

```{r setup}
library(tidyverse)
library(Seurat)
library(ggplot2)
library(cowplot)
library(yardstick)
library(DropletUtils)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("functions/plotting_fxns.R")
source("functions/scRNA_seq_analysis_functions.R")
theme_set(theme_Publication())
```

### [Import Data]{.underline}

```{r}
# dataCD4 <- readRDS("/Users/wange7/Library/CloudStorage/GoogleDrive-ericwang314@gmail.com/My Drive/Lab/datasets/EYW/SIG03_10x_240706/seurat_outs/SIG03_mouse_oBC_DirectAmp_CD4.rds")

dataCD4 <- readRDS("C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/seurat_outs/SIG03_mouse_oBC_DirectAmp_CD4.rds")
```

### [Downsample Data]{.underline}

This library had 592,600,865 reads with a sequencing saturation of 72%. Within the library, 75.3% of reads were usable resulting in a final count of 8,857 reads per cell.

Targeting 500 reads per cell would correspond to 30 million reads or 5% downsample

```{r}
# downsample and add rownames
downBC500 <- downsampleReads("C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/cellranger_outs/SIG03_GEX_HTO_oBC_DirectAmp_mouse/molecule_info.h5",
                           0.05, bycol = FALSE, use.library = "Custom")
downBC1000 <- downsampleReads("C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/cellranger_outs/SIG03_GEX_HTO_oBC_DirectAmp_mouse/molecule_info.h5",
                           0.05, bycol = FALSE, use.library = "Custom")
downBC2000 <- downsampleReads("C:/Users/Eric/My Drive/Lab/datasets/EYW/SIG03_10x_240706/cellranger_outs/SIG03_GEX_HTO_oBC_DirectAmp_mouse/molecule_info.h5",
                           0.05, bycol = FALSE, use.library = "Custom")

# subset to only include cell bc present in processed data
downBC500 <- downBC500[,colnames(dataCD4)]
downBC1000 <- downBC1000[,colnames(dataCD4)]
downBC2000 <- downBC2000[,colnames(dataCD4)]
```

### [Visualize Barcode Distributions]{.underline}

```{r, fig.width=10, fig.height=10}
expressionData500 <- tibble(cell_bc = colnames(dataCD4),
                         hash.ID = dataCD4$hash.ID,
                         BC4 = downBC500["p139-BC5",],
                         BC5 = downBC500["p139-BC4",])
expressionData1000 <- tibble(cell_bc = colnames(dataCD4),
                         hash.ID = dataCD4$hash.ID,
                         BC4 = downBC1000["p139-BC5",],
                         BC5 = downBC1000["p139-BC4",])
expressionData2000 <- tibble(cell_bc = colnames(dataCD4),
                         hash.ID = dataCD4$hash.ID,
                         BC4 = downBC2000["p139-BC5",],
                         BC5 = downBC2000["p139-BC4",])
expressionData <- tibble(cell_bc = colnames(dataCD4),
                         hash.ID = dataCD4$hash.ID,
                         BC4 = dataCD4@assays$BC$counts["p139-BC5",],
                         BC5 = dataCD4@assays$BC$counts["p139-BC4",])

p1 <- expressionData500 %>%
  filter(hash.ID %in% c("CD4-6h-2e10")) %>%
  ggplot(aes(x = log10(BC4 + 1), y = log10(BC5 + 1))) +
    geom_point() +
    stat_density2d(aes(fill = after_stat(level)), geom = "polygon") +
    scale_fill_viridis_c() +
    theme(aspect.ratio = 1) +
    xlim(0,4.3)+
    ylim(0,4.3) +
    ggtitle("6h 2e10 LP/mL 500 reads/cell")

p2 <- expressionData1000 %>%
  filter(hash.ID %in% c("CD4-6h-2e10")) %>%
  ggplot(aes(x = log10(BC4 + 1), y = log10(BC5 + 1))) +
    geom_point() +
    stat_density2d(aes(fill = after_stat(level)), geom = "polygon") +
    scale_fill_viridis_c() +
    theme(aspect.ratio = 1) +
    xlim(0,4.3)+
    ylim(0,4.3) +
    ggtitle("6h 2e10 LP/mL 1000 reads/cell")

p3 <- expressionData2000 %>%
  filter(hash.ID %in% c("CD4-6h-2e10")) %>%
  ggplot(aes(x = log10(BC4 + 1), y = log10(BC5 + 1))) +
    geom_point() +
    stat_density2d(aes(fill = after_stat(level)), geom = "polygon") +
    scale_fill_viridis_c() +
    theme(aspect.ratio = 1) +
    xlim(0,4.3)+
    ylim(0,4.3) +
    ggtitle("6h 2e10 LP/mL 2000 reads/cell")

p4 <- expressionData %>%
  filter(hash.ID %in% c("CD4-6h-2e10")) %>%
  ggplot(aes(x = log10(BC4 + 1), y = log10(BC5 + 1))) +
    geom_point() +
    stat_density2d(aes(fill = after_stat(level)), geom = "polygon") +
    scale_fill_viridis_c() +
    theme(aspect.ratio = 1) +
    xlim(0,4.3)+
    ylim(0,4.3) +
    ggtitle("6h 2e10 LP/mL full")

plot_grid(p1,p2,p3,p4, ncol = 2)
```

### [Barcode Classification]{.underline}

```{r}
data6hCD4 <- subset(dataCD4, subset = hash.ID %in% c("CD4-6h-2e10","CD4-6h-0"))

downBC5006h <- downBC500[,colnames(data6hCD4)]
downBC10006h <- downBC1000[,colnames(data6hCD4)]
downBC20006h <- downBC2000[,colnames(data6hCD4)]
```

#### Barcode 4 Recall

Create test datasets for 6h

```{r}
testDataList <- vector(mode = "list")

testDataList[["500"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC5006h["p139-BC4",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["1000"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC10006h["p139-BC4",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["2000"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC20006h["p139-BC4",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["full"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = data6hCD4@assays$BC$counts["p139-BC4",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))
```

```{r}
# Function to classify cells based on UMI count cutoff
classify_cells <- function(data, umi_cutoff) {
  data %>%
    mutate(predicted_label = ifelse(umi_counts >= umi_cutoff, "positive", "negative")) %>%
    mutate(predicted_label = factor(predicted_label, c("positive","negative")))
}

# range of UMI cutoffs to test
umi_cutoffs <- seq(0,1000, by = 1)

results <- tibble()
for(i in 1:length(testDataList)){
  temp <- map_df(umi_cutoffs, function(cutoff) {
  classified_data <- classify_cells(testDataList[[i]], cutoff)
  
  metrics <- classified_data %>%
    precision(truth = label, estimate = predicted_label) %>%
    bind_rows(
      classified_data %>%
        recall(truth = label, estimate = predicted_label),
      classified_data %>%
        sensitivity(truth = label, estimate = predicted_label),
      classified_data %>%
        specificity(truth = label, estimate = predicted_label),
    )
  metrics %>%
    mutate(umi_cutoff = cutoff)
  })
  # add category label to results
  temp <- mutate(temp, group = names(testDataList)[i])
  results <- bind_rows(results,temp)
}
write_csv(results, "analysis_outs/umi_cutoff_oBCDirect_metrics_BC4_6h_2e10_downsample_cell.csv")

# plot PR curve
pr1 <- results %>%
  filter(.metric %in% c("precision", "recall")) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(group = factor(group, c("500","1000","2000","full"))) %>%
  ggplot(aes(x = recall, y = precision, color = group)) +
    geom_line(linewidth=1) +
    scale_color_brewer(palette = "Dark2") +
    labs(title = "PR Curve 6h BC4", x = "Recall", y = "Precision") +
    theme(aspect.ratio = 1) +
    xlim(0.5,1) +
    ylim(0.5,1)

# plot ROC
roc1 <- results %>%
  filter(.metric %in% c("sensitivity", "specificity")) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(FPR = 1-specificity) %>%
  mutate(group = factor(group, c("500","1000","2000","full"))) %>%
  ggplot(aes(x = FPR, y = sensitivity, color = group)) +
    geom_line(linewidth=1) +
    scale_color_brewer(palette = "Dark2") +
    labs(title = "ROC 6h BC4", x = "False Positive Rate", y = "True Positive Rate") +
    theme(aspect.ratio = 1) +
    xlim(0,1) +
    ylim(0,1)
```

#### Barcode 5 Recall

Create test datasets for 6h

```{r}
testDataList <- vector(mode = "list")

testDataList[["500"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC5006h["p139-BC5",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["1000"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC10006h["p139-BC5",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["2000"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = downBC20006h["p139-BC5",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))

testDataList[["full"]] <- tibble(hash.ID = data6hCD4$hash.ID,
                   umi_counts = data6hCD4@assays$BC$counts["p139-BC5",]) %>%
  filter(hash.ID %in% c("CD4-6h-2e10","CD4-6h-0")) %>%
  mutate(label = case_when(hash.ID == "CD4-6h-0" ~ "negative",
                           hash.ID == "CD4-6h-2e10" ~ "positive")) %>%
  mutate(label = factor(label, c("positive","negative")))
```

```{r}
# Function to classify cells based on UMI count cutoff
classify_cells <- function(data, umi_cutoff) {
  data %>%
    mutate(predicted_label = ifelse(umi_counts >= umi_cutoff, "positive", "negative")) %>%
    mutate(predicted_label = factor(predicted_label, c("positive","negative")))
}

# range of UMI cutoffs to test
umi_cutoffs <- seq(0,1000, by = 1)

results <- tibble()
for(i in 1:length(testDataList)){
  temp <- map_df(umi_cutoffs, function(cutoff) {
  classified_data <- classify_cells(testDataList[[i]], cutoff)
  
  metrics <- classified_data %>%
    precision(truth = label, estimate = predicted_label) %>%
    bind_rows(
      classified_data %>%
        recall(truth = label, estimate = predicted_label),
      classified_data %>%
        sensitivity(truth = label, estimate = predicted_label),
      classified_data %>%
        specificity(truth = label, estimate = predicted_label),
    )
  metrics %>%
    mutate(umi_cutoff = cutoff)
  })
  # add category label to results
  temp <- mutate(temp, group = names(testDataList)[i])
  results <- bind_rows(results,temp)
}
write_csv(results, "analysis_outs/umi_cutoff_oBCDirect_metrics_BC5_6h_2e10_downsample_cell.csv")

# plot PR curve
pr3 <- results %>%
  filter(.metric %in% c("precision", "recall")) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(group = factor(group, c("500","1000","2000","full"))) %>%
  ggplot(aes(x = recall, y = precision, color = group)) +
    geom_line(linewidth=1) +
    scale_color_brewer(palette = "Dark2") +
    labs(title = "PR Curve 6h BC5", x = "Recall", y = "Precision") +
    theme(aspect.ratio = 1) +
    xlim(0.5,1) +
    ylim(0.5,1)

# plot ROC
roc3 <- results %>%
  filter(.metric %in% c("sensitivity", "specificity")) %>%
  pivot_wider(names_from = .metric, values_from = .estimate) %>%
  mutate(FPR = 1-specificity) %>%
  mutate(group = factor(group, c("500","1000","2000","full"))) %>%
  ggplot(aes(x = FPR, y = sensitivity, color = group)) +
    geom_line(linewidth=1) +
    scale_color_brewer(palette = "Dark2") +
    labs(title = "ROC 6h BC5", x = "False Positive Rate", y = "True Positive Rate") +
    theme(aspect.ratio = 1) +
    xlim(0,1) +
    ylim(0,1)
```


#### Visualize

```{r, fig.height=12, fig.width=12}
plot_grid(pr1,pr3,roc1,roc3)
```



